Iroko: Predictive Datacenter Congestion Control

### 2017-12-11 Documents/Report/implementation.tex
Implementation updates

\subsection{Topology Simulation}
We chose to use  Mininet, a python data center emulation API, to prototype Iroko. Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}. This is an important advantage, because it allowed us to create a prototype we can test using limited amounts of hardware. Other methods of testing Iroko would require more resources, or require specialized code specific to the testbed which cannot deployed in a real system ~\cite{mininet}. In contrast, Mininet has allowed us to develop Iroko to be deployable in a real system without re-implementation independent of testbed specific code.

Mininet’s lightweight virtualization also allows the API to scale better, and to support bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}. This feature will be useful for future research, because we can then test Iroko on other topologies beyond the fat-tree topology we looked at in this work.

In the Mininet environment, we created a fat-tree topology consisting of 20 switches and 16 hosts . The 20 switches are configured into 5 pods as described in ~\cite{fattree}. This design choice was to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics to re-route elephant flows. In ~\cite{hedera}, Hedera was evaluted  using a fat-tree topology as we have described, and this allows us to directly compare Iroko against Hedera's performance .

\subsection{Data Collection }
We collected several different metrics from the links connected to each host. We leverage the fact we know which host connects to which link in the data center. We used the awk scripting language and the Linux tc qdisc command to collect information from the links.

Awk is a text parsing language available on most Linux distributions by default. The awk script reads the /proc/net/dev file to collect the currently used  bandwidth on a link. We then used this to calculate free bandwidth which is the difference between theoretical maximum bisection bandwidth, and the reported bandwidth on a link.

In addition, we used the following Linux command:
\[$tc -s qdisc show dev <interface>$ \] 
where $<interface>$ is the current host we are polling from. Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues . Each of these data provide further information to hint at the current state of the data center. Our goal was to eliminate dropped packets, and used this to help Iroko make rate-limiting decisions. Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from. Interface queues provide Iroko with information on the queue length of the host.  These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.

@@

In order to make rate-limiting decisions per host, we treat bandwidth allocation as a reinforcement learning problem. The goal in reinforcement learning is to maximize the expected reward given by an environment for each possible state $s$ ~\cite{Sutton:1998:IRL:551283}. These values are often tracked using the state-action value function $Q(s,a)$, and often this value is approximated using a function due to intractable memory needs ~\cite{Sutton:1998:IRL:551283}. Using $Q(s,a)$, a policy $\pi(s)$ is then learned to direct how the agent should proceed in the environment.$\pi(s)$ role is to decide the next action a reinforcement learning agent should take, in our case the amount to rate-limit a host.

@@

The action Iroko takes at each step is the amount to increase or decrease the current hosts allocated bandwidth. Iroko will generate an adjustment value clamped between $-.9 - .9$. We chose this value range, because we wanted to avoid starving hosts.  Iroko then modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:

>>>>

\subsection{Topology Simulation}
We chose to use  Mininet, a python data center emulation API, to prototype Iroko. Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}. This allowed us to create a prototype of Iroko we can test quickly using limited amounts of hardware. Other methods of testing Iroko would require more resources, or specialized code specific to the testbed which cannot deployed in a real system ~\cite{mininet}. In contrast, Mininet has allowed us to develop Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.

In the Mininet environment, we created a fat-tree topology consisting of 20 switches and 16 hosts . The 20 switches are configured into 5 pods as described in ~\cite{fattree}. This design choice was to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics to re-route elephant flows. In ~\cite{hedera}, Hedera was evaluated  using a fat-tree topology as we have described, and this allows us to directly compare Iroko against Hedera's performance .

Mininet’s lightweight virtualization also allows the API to scale such as to support bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}. This feature will be useful for future research, because we can then test Iroko on other topologies beyond the fat-tree topology we looked at in this work.

\subsection{Data Collection }
Since we knew prior to testing which port connected to each host, we used different link statistics from each switch port connected to host to make predictions. This was accomplished using the awk scripting language and the Linux tc qdisc command to collect information.

Awk is a text parsing language available on most Linux distributions by default. Our awk script reads the /proc/net/dev file on each interface to collect the currently used  bandwidth on a link. We then used this to calculate free bandwidth which is the difference between theoretical maximum bisection bandwidth, and the reported bandwidth on a link.

In addition, we used the following Linux command:
\[$tc -s qdisc show dev <interface>$ \] 
where $<interface>$ is the current host we are polling from. Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues. Each of these data provide further information to hint at the current state of the data center. Our goal was to eliminate dropped packets, and used this to help Iroko make rate-limiting decisions. Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from. Interface queues provide Iroko with information on the queue length of the host, which measures how many packets are buffered to be processed.  These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.

@@

In order to make rate-limiting decisions per host, we treat bandwidth allocation as a reinforcement learning problem. The goal in reinforcement learning is to maximize the expected reward given by an environment for each possible state $s$ by choosing the best action $a$ to take at each time step~\cite{Sutton:1998:IRL:551283}. These values are can be tracked using the state-action value function $Q(s,a)$, and often this value is approximated using a function due to intractable memory needs ~\cite{Sutton:1998:IRL:551283}. Using $Q(s,a)$, a policy $\pi(s)$ is then learned to direct how the agent should proceed in the environment.$\pi(s)$ role is to decide the next action a reinforcement learning agent should take, in our case the amount to rate-limit a host.

@@

The action Iroko takes at each step is the amount to increase or decrease the current hosts allocated bandwidth. Iroko will generate an adjustment value clamped between $-.9 - .9$. We chose this value range, because we wanted to avoid host starvation.  Iroko then modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:
